---
title: "MA415 Midterm Project"
author: "Rachel Feingold"
date: "3/7/2017"
output: pdf_document
---
Instructions:

Prepare the OSHA data and create a dataset for analysis.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Libraries Needed:
```{r}
require(foreign)
require(dplyr)
require(tidyr)
require(ggplot2)
require(lubridate)
require(stringr)
require(magrittr)
require(data.table)
```


Reading Data:
```{r}
osha <- read.dbf("osha.dbf")
# accidents <- read.dbf("accid.dbf")
# violations <- read.dbf("viol.dbf")
# hazsub <- read.dbf("hazsub.dbf")
# history <- read.dbf("history.dbf")
# admpay <- read.dbf("admpay.dbf")

```

Cleaning Data

1. First, I cleaned the OSHA table. I am starting with this one because it is the main table and it has the company names, address, date of inspection, and other descriptive information about the companies.

2. Then, I selected the variables (columns) I deemed most important to the analysis of the most dangerous places to work in Massachusetts. I dropped the rest. 
```{r} 
tidyosha <- osha[-c(1:4, 9:11, 14:15, 17, 19:20, 22:24, 27:33, 35:59, 61:63, 68:77, 80:92) ]
```
3. Then, I checked to see which variables were stable, meaning that the value of the variables don't change at all. I could discard 6 variables. However, we want to extract the stable information of the data set just in case. Afterwards, I dropped the stable variables. 
```{r}
table(tidyosha$OPTREPTNO) #discard
table(tidyosha$SITESTATE) #discard
table(tidyosha$EMPCOUNT) #discard
table(tidyosha$EMPCOVERED) #discard
table(tidyosha$NATEMPCNT)#discard
table(tidyosha$LWDIRATE) #discard

StableInfo = tidyosha[,c("OPTREPTNO","SITESTATE", "EMPCOUNT", "EMPCOVERED", "NATEMPCNT", "LWDIRATE")][1,]

dr = function(name){return(which(colnames(tidyosha)==name))}
tidyosha = tidyosha[,-c(dr("OPTREPTNO"),
                        dr("SITESTATE"),
                        dr("EMPCOUNT"),
                        dr("EMPCOVERED"),
                        dr("NATEMPCNT"),
                        dr("LWDIRATE"))]
rm(i,dr)
```
4. Then, I chose from the remaining variables, which would be the most useful for analysis. 
```{r}
tidyosha =tidyosha[,c(dr(""))]

```
5. Next, I cleaned accidents. From scanning the database in the view function, there are no columns with all NA data. So I just looked for stable variables.
```{r} 
table(accidents$SITESTATE) #discard
# table(accidents$OCC_CODE)

dr = function(name){return(which(colnames(accidents)==name))}
accidents = accidents[,-c(dr("SITESTATE"))]
               
rm(i,dr)
```
Organizing Data